---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggridges)
library(patchwork)
library(ggplot2)


Sys.setenv("plotly_username"="theonlyviper66")
Sys.setenv("plotly_api_key"="RXpSTZ2TUHpbh6D0Nr8r")
Sys.setenv('MAPBOX_TOKEN' = 'pk.eyJ1IjoidGhlb25seXZpcGVyNjYiLCJhIjoiY2xiZWY3dmp6MGFyeTNwcXppdzRnYm1nbyJ9.G05d_geqdk2HTTdZRmOmzg')

```

# Restaurant Review 

```{r include=FALSE}
inspection_raw<-read.csv("data/inspection_sub_latest_date.csv") %>% 
  janitor::clean_names()
```

## Top 10 cuisine in the whole NYC and by boro
As a real international metropolitan, There is a total of 80 different cuisine types in NYC based on our dataset and we firstly plot the 10 most frequently shown cuisine types in NYC.
```{r echo = FALSE, warning=FALSE}
top_10_cuisine_nyc<-
  inspection_raw %>% 
  drop_na(boro) %>%
  group_by(cuisine_description) %>% 
  summarise(n=n()) %>% 
  filter(min_rank(desc(n))<=10) %>%
  mutate(cuisine_description=fct_reorder(cuisine_description,n,.desc=TRUE)) %>% 
  ggplot(aes(x=cuisine_description,y=n,fill=cuisine_description))+
  geom_bar(position="dodge", stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  labs(
    x="cuisine type",
    y="Number"
  )

ggplotly(top_10_cuisine_nyc)
```

From the plot above, the top 10 cuisine types across NYC include `American`, `Chinese`, `Coffee/Tea`, `Pizza`, `Bakery Products/Desserts`, `Mexican`, `Japanese`, `Italian`, `latin American`, `Caribbean`, which implies the racial diversity of NYC. `American`, `Chinese`, `Coffee/Tea` are the top 3 favorite cuisine types for NY citizens. We then inspect the top 10 frequently shown cuisine types in different boroughs. 

```{r echo = FALSE, warning=FALSE}
boro_cuisine_bar<-
inspection_raw %>% 
  select(dba,boro,cuisine_description,critical_flag,score,grade,grade_date,inspection_type,latitude,longitude,rating,review_num,price) %>%
  drop_na(boro) %>% 
  mutate(cuisine_description=fct_infreq(cuisine_description),
         boro=fct_infreq(boro)) %>% 
  group_by(boro,cuisine_description) %>%
  summarize(n_obs=n()) %>% 
  filter(min_rank(desc(n_obs)) <= 10) %>%
  ggplot(aes(x=cuisine_description,y=n_obs,fill=cuisine_description))+
  geom_bar(position="dodge", stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  theme(axis.text.x = element_text(size = 5))+
  labs(
    x="Cuisine type",
    y="Number")+
  facet_grid(~boro)
ggplotly(boro_cuisine_bar)
```
The top 3 frequently cuisine types in `Manhattan` and `Brooklyn` are `American`, `Coffee/Tea`, and `Chinese`. 


## price vs cuisine

```{r}
price_cuisine<-
  inspection_raw %>% 
  select(dba,boro,cuisine_description,critical_flag,score,grade,grade_date,inspection_type,latitude,longitude,rating,review_num,price) %>%
  drop_na(boro,price) %>% 
  mutate(price=as.factor(price)) %>% 
  group_by(price)%>%
  count(cuisine_description) %>%
  mutate(cuisine_description=fct_reorder(cuisine_description,n)) %>%
  filter(min_rank(desc(n))<=10) %>%
  ggplot(aes(x=cuisine_description,y=n,fill=price))+
  geom_bar(position="dodge", stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  theme(axis.text.x = element_text(size = 5))+
  labs(
    x="Cuisine type",
    y="Number")+
  facet_grid(~price)
ggplotly(price_cuisine)
```

## Price vs. borough

```{r}
price_boro<-
  inspection_raw %>% 
  select(dba,boro,cuisine_description,critical_flag,score,grade,grade_date,inspection_type,latitude,longitude,rating,review_num,price) %>%
  drop_na(boro,price) %>%
  mutate(boro = fct_infreq(boro),
         price=as.factor(price)) %>%
  ggplot(aes(x = boro, fill = price)) + 
  geom_bar()
ggplotly(price_boro)
```

## review number vs. rating score (exclude extreme outliers)

```{r}
review_num_rating<-
  inspection_raw %>% 
  select(dba,boro,cuisine_description,critical_flag,score,grade,grade_date,inspection_type,latitude,longitude,rating,review_num,price) %>% drop_na(boro,price,rating) %>%
  filter(!review_num<=100) %>%
  group_by(rating) %>%
  summarize(sum_review=sum(review_num)) %>% 
  ggplot(aes(y=sum_review,x=rating))+
  geom_point()+
  geom_smooth()+
  labs(
    x="Sum of review_numbers",
    y="Rating"
  )
ggplotly(review_num_rating)
```


## Review Scores & Price  

```{r}
inspection_raw %>% 
  filter(!is.na(price)) %>%
  ggplot(aes(x = price, y = rating, fill = price), data = .)+geom_boxplot()+labs(title = "Yelp Rating vs. Price")

inspection_raw %>% 
  filter(!is.na(price)) %>%
  ggplot(aes(y = rating, fill = price), data = .)+geom_density()+labs(title = "Density Plot")+facet_grid(~price)
```

There seems to be a positive relationship between the cost of dining and yelp rating of NYC restaurants. The distribution of restaurant review scores appears to be right-skewed, probably due to the presence of outliers with a much lower value compared to the majority of the data. 

## Geolocation of Restaurants by Price 

This map displays the geographical locations of the restaurants in the dataset. The map is interactive and is split on the costs of dining.   

```{r}
price_rating_map = inspection_raw %>%
  mutate(text_label = str_c("Price: ", price, "\nRating: ", rating)) %>% 
  drop_na(price) %>%
  plot_mapbox(
    lat = ~latitude,
    lon = ~longitude,
    mode = "markers",
    split = ~price,
    mode = "markers",
    hovertext = ~text_label) %>%
  layout(
    mapbox = list(
      style = 'dark',
      zoom =12.5,
      center = list(lat = 40.71, lon = -73.98))) 
price_rating_map %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))
```

In general, Manhattan has the largest number of restaurants and a much denser distribution compared to other boroughs. The proportion of the least and second least expensive restaurants are much higher compared to the proportions of more expensive dining places in Bronx, Queens, Brooklyn, and Staten Island. In addition, the majority of the restaurants that fall into the most expensive category are located in Manhattan.  

## cuisine map

```{r}
price_rating_map = inspection_raw %>%
  mutate(text_label = str_c("Cuisine", cuisine_description)) %>% 
  drop_na(price) %>%
  plot_mapbox(
    lat = ~latitude,
    lon = ~longitude,
    mode = "markers",
    split = ~cuisine_description,
    mode = "markers",
    hovertext = ~text_label) %>%
  layout(
    mapbox = list(
      style = 'dark',
      zoom =12.5,
      center = list(lat = 40.71, lon = -73.98))) 
price_rating_map %>% config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))
```
