---
title: "Statistical Analysis"
author: "Jiayi Shi"
date: "2022-12-09"
Output:
  html_notebook:
    code_folding: hide
---
```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(gtsummary)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r}
library(readr)
dt = read_csv("data/inspection_sub_all_date.csv")
df = read_csv("data/inspection_sub_latest_date.csv") %>%
  distinct(dba, .keep_all = TRUE) %>% 
  mutate(price = fct_recode(price, "1" = "$", "2" = "$$", "3" = "$$$", "4" = "$$$$"),
         boro = fct_reorder(boro, rating),
         grade = fct_relevel(grade, "A"))

```

```{r}
df %>% 
  select(boro, grade, price) %>% 
  mutate(boro = fct_infreq(boro)) %>% 
  tbl_summary(
    by = boro,
    missing_text = "(Missing)", 
    statistic = list(
      all_categorical() ~ "{n} ({p}%)"
      )) %>% 
  bold_labels() %>%   
  italicize_levels() 

df %>% 
  select(score,rating, review_num) %>% 
  tbl_summary(
    missing_text = "(Missing)", 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
      )) %>% 
  bold_labels() %>%   
  italicize_levels() 
```

## Chi-square Test

We try to determine whether there is a relationship between boroughs and restaurants' inspection grades. Our hypothesis is that there is no difference in the number of restaurants across the five grades across the five boroughs in NYC. We will perform the chi-square test to verify our assumption.

$H0$: the expected number of restaurants in each grades are the same across all boroughs.

$H1$: the expected number of restaurants in each grades are not same across all boroughs.

```{r}
grade_boro = 
  df %>% 
  drop_na(grade) %>% 
  count(boro, grade) %>% 
  pivot_wider(
    names_from = "grade",
    values_from = "n") %>%
  replace(is.na(.), 0) %>% 
  data.matrix() %>%
  subset(select = -c(boro))

rownames(grade_boro) <- c("Bronx", "Staten Island", "Queens", "Brooklyn", "Manhattan")

grade_boro %>% 
  knitr::kable(caption = "Results Table")
```

```{r}
chisq.test(grade_boro)
```

Interpretation: The result of chi-square shows that p-value is less than 0.05, so we reject the null hypothesis at 95% significant level and conclude that the inspection grades of restaurants are significantly different by boroughs. 

```{r}
price_grade = df %>% 
  select(boro, score, grade, rating, review_num, price) %>%
  drop_na(price, grade) %>% 
  group_by(price, grade) %>% 
  summarise(n = n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n 
  ) %>% 
  replace(is.na(.), 0) %>% 
  data.matrix() %>%
  subset(select = -c(price))

rownames(price_grade) <- c("$", "$$", "$$$", "$$$$")

price_grade %>% 
  knitr::kable(caption = "Results Table")

chisq.test(price_grade)
```

Interpretation: The result of chi-square shows that p-value is less than 0.05, so we reject the null hypothesis at 95% significant level and conclude that the inspection grades of restaurants are significantly different by price scales of restaurants.


## Proportion Test

Now, we want to see whether receiving grade A is equally common among restaurants of all four price scales. To do this, we will conduct a proportion test.

```{r}
total = df %>% 
  group_by(price) %>% 
  summarise(total = n())

n_a = df %>% 
  count(price, grade) %>% 
  filter(grade == "A")

join = left_join(total, n_a) %>% drop_na()
prop.test(join$n, join$total)

#join = join %>% 
 # mutate(prop = n/total) %>% 
 # select(price, prop)
```

From the test result, we can see that the p-value is greater than 0.05,  

## Regression model

* Response Transformation
```{r}
df %>% 
  select(boro, cuisine_description, score, grade, rating, review_num, price) %>% 
  ggplot(aes(x = rating))+
  geom_histogram(aes(y = ..density..), bins = 30, color = "black", fill = "grey") +
  geom_density()

qqnorm(df$rating)
```

Although the response variable `rating` is not normally distributed, linear regression remains a statistically sound technique since the sample size is large enough (i.e.>3000).

* Predictor transfomation

```{r}
top_cuisine = df %>%
  count(cuisine_description) %>% 
  arrange(-n) %>% 
  top_n(4) %>% 
  pull(cuisine_description)

reg_df = df %>% 
  select(boro, cuisine_description, score, grade, rating, review_num, price) %>%
  filter(cuisine_description %in% top_cuisine) %>% 
  filter(grade %in% c("A", "B", "C"))
  

dlookr::find_skewness(reg_df)

par(mfrow = c(2, 2))
plot(rating ~ review_num, data = reg_df, col = "dodgerblue", pch = 20, cex = 1.5)
review = lm(rating ~ review_num, data = reg_df)
abline(review, col = "darkorange", lwd = 2)
plot(fitted(review), resid(review), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
title(main = "Before transformation")

plot(rating ~ log(review_num), data = reg_df, col = "dodgerblue", pch = 20, cex = 1.5)
review = lm(rating ~ log(review_num), data = reg_df)
abline(review, col = "darkorange", lwd = 2)
plot(fitted(review), resid(review), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
title(main = "After transformation")

plot(rating ~ score, data = reg_df, col = "dodgerblue", pch = 20, cex = 1.5)
rating_score = lm(rating ~ score, data = reg_df)
abline(rating_score, col = "darkorange", lwd = 2)
plot(fitted(rating_score), resid(rating_score), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
```

Log transform review numbers:

```{r}
reg_df = reg_df %>% 
  select(rating, everything()) %>% 
  mutate(review_num = log(review_num))
```

Rating has some association with cuisine type, inspection grade and price scale.

```{r}
reg_df %>% 
  ggplot(aes(y = rating, color = cuisine_description))+
  geom_boxplot()

reg_df %>%
  ggplot(aes(y = rating, color = grade))+
  geom_boxplot()

reg_df %>%
  drop_na(price) %>% 
  ggplot(aes(y = rating, color = price))+
  geom_boxplot()

reg_df %>%
  mutate(boro = fct_reorder(boro, rating)) %>% 
  ggplot(aes(y = rating, color = boro))+
  geom_boxplot()
```

There is interaction between score and cuisine type:
```{r}
reg_df %>% 
  ggplot(aes(x = score, y = rating, color = cuisine_description))+
  geom_point()+
  geom_smooth(method="lm", se=F, aes(group = cuisine_description, color = cuisine_description))
```

Score has little influence on rating, borough is not a confounder.
```{r}
reg_df %>% 
  ggplot(aes(x = score, y = rating, color = boro))+
  geom_point()+
  geom_smooth(method="lm", se=F, aes(group = boro, color = boro))
```


```{r}
model = lm(rating ~ cuisine_description*score+boro+review_num+price, data = reg_df)
summary(model)
```

10-fold cross validation:
```{r}
library(caret)
set.seed(1)
train = trainControl(method = "cv", number = 10)
model = train(rating ~ cuisine_description*score+boro+review_num+price, 
              data = reg_df,
              trControl = train,
              method = 'lm',
              na.action = na.pass)
model$finalModel
print(model)
```

Elastic Net
```{r}
library(glmnet)
library(caret)
set.seed(2)
cv_10 = trainControl(method = "cv", number = 10)
x = model.matrix(~ ., dplyr::select(reg_df, -rating))[,-1]
y = reg_df$rating
elnet_int = train(
  rating ~ . ^ 2, data = reg_df %>% drop_na(),
  method = "glmnet",
  trControl = cv_10,
  tuneLength = 10
)

get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}

get_best_result(elnet_int)
```

