---
title: "Statistical Analysis"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(gtsummary)

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  fig.width = 6,
  fig.asp = .8,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



## Dataset Overview

### Data Wrangling

In this first step, we wrangle the data such that it is suitable for statistical analysis and provide an overview for the dataset of use. In particular, we create a new numeric variable, percent_critical, that accounts for the percentage of critical flags a given restaurant receives out of all violations. Moreover, we factorize categorical variables price, borough, and grade; before finally making distinct each restaurant such that each represent a unique row. 

```{r}
library(readr)
# dt = read_csv("data/inspection_sub_all_date.csv")
df = read_csv("data/inspection_sub_latest_date.csv") %>%
  mutate(price = fct_recode(price, "1" = "$", "2" = "$$", "3" = "$$$", "4" = "$$$$"),
         boro = fct_reorder(boro, rating),
         grade = fct_relevel(grade, "A")) %>%
  group_by(camis) %>%
  mutate(percent_critical = sum(critical_flag == "Critical")/n())

# Make distinction such that each restaurant is its unique row.
df <- df %>% 
  distinct(camis, .keep_all = TRUE) %>%
  ungroup() %>% # select desired variables
  select(dba, boro, grade, zipcode, cuisine_description, inspection_type, rating, review_num, price, percent_critical, score)

# Display data
head(df)
```


### Summary of Wrangled Data

Now, a brief summary of our dataset to use.

```{r}
df %>% 
  ungroup() %>%
  select(boro, grade, price) %>% 
  mutate(boro = fct_infreq(boro)) %>% 
  tbl_summary(
    by = boro,
    missing_text = "(Missing)", 
    statistic = list(
      all_categorical() ~ "{n} ({p}%)"
      )) %>% 
  bold_labels() %>%   
  italicize_levels() 

df %>% 
  ungroup() %>%
  select(score,rating, review_num) %>% 
  tbl_summary(
    missing_text = "(Missing)", 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
      )) %>% 
  bold_labels() %>%   
  italicize_levels() 
```


From these results, we observe that: 

* Manhattan has the most restaurants (40% of all restaurants in the city), followed by Brooklyn (28%), Queens (22%), Bronx (6.6%), and Staten Island (3.4%).

* The "score" variable has 100 missing values with mean 18 and standard deviation 14.

* The "grade" variable has 1,399 missing values with most of the restaurants receiving As (70% of all restaurants).

* Each restaurant has average rating 3.78 with standard deviation 0.68.

* Each restaurant has on average 232 reviews with standard deviation 485.

* The most common price range for restaurants is between \$11-\$30 (60%), followed by under \$10 (31%).



## Chi-square Test

Now, we conduct chi-square tests to get a sense of whether boro and price range has an effect on the inspection grade that a given restaurant receives. Since the inspection grades reflect the extent of health hazards one might take risk of in dining at that particular restaurant and has been required by the City since 2010 for restaurants to post letter grades that correspond to scores received from sanitary inspections, this investigation might shed light on the whether dining at a particular borough or price range increases one's danger of food poisoning or other food-borne diseases.


### Inspection Grades and Boroughs

We try to determine whether there is a relationship between boroughs and restaurants' inspection grades. Our hypothesis is that there is no difference in the number of restaurants across the five grades across the five boroughs in NYC. We will perform the chi-square test to verify our assumption.

$H0$: the expected number of restaurants in each grades are the same across all boroughs.

$H1$: the expected number of restaurants in each grades are not same across all boroughs.

```{r}
grade_boro = 
  df %>% 
  ungroup() %>%
  filter(grade %in% c("A", "B", "C")) %>% 
  count(boro, grade) %>% 
  pivot_wider(
    names_from = "grade",
    values_from = "n") %>%
  replace(is.na(.), 0) 

# Display results table.
grade_boro %>% 
  knitr::kable(caption = "Results Table",
               col.names = c("Borough", "A", "B", "C"))

grade_boro = grade_boro %>%  
  data.matrix() %>%
  subset(select = -c(boro))

# Perform chi-square test
chisq.test(grade_boro)
```

Interpretation: Since p-value is less than 0.05, we reject the null hypothesis at 95% significant level and conclude that the inspection grades of restaurants are significantly different by boroughs. In other words, dining at different boroughs in NYC carries different risk of food-related health hazards.


### Price and Inspection Grade

Now, the reason that many might choose a restaurant of higher price range out of safety concerns. In this part we investigate if such is true with restaurants in NYC: does price range of a restaurant has an effect on its safety levels?

$H0$: There is no difference in the inspection grade received among the four different groups of price range.

$H1$: There is a difference in the inspection grade received among the four different groups of price range.

```{r}
price_grade = df %>% 
  select(boro, score, grade, rating, review_num, price) %>%
  drop_na(price, grade) %>% 
  filter(grade %in% c("A", "B", "C")) %>% 
  group_by(price, grade) %>% 
  summarise(n = n()) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n 
  ) %>% 
  replace(is.na(.), 0) %>% 
  mutate(price = fct_recode(price, "<10"="1", "11-30"="2", "31-60"="3", ">60"="4"))

price_grade %>% 
  knitr::kable(caption = "Results Table",
               col.names = c("Price($)", "A", "B", "C"))

price_grade = price_grade %>% 
  data.matrix() %>%
  subset(select = -c(price))

# Perform chi-square test
chisq.test(price_grade)
```

Interpretation: The result of chi-square shows that p-value is more than 0.05, so we fail to reject the null hypothesis at 95% significant level and conclude that we don't have sufficient evidence to claim that the inspection grades of restaurants differ by price scales of restaurants. In other words, the results from this analysis tells us that the urban myth that dining in more expensive restaurant avoids health hazards is unfounded.


## Proportion Test


### Price and Receiving "A"

Now, we want to see whether receiving grade A is equally common among restaurants of all four price scales. To do this, we will conduct a proportion test.

$H0$: Receiving A is equally common among restaurants fo all four price scales.

$H1$: Receiving A is not equally common among restaurants for all four price scales.

```{r}
total = df %>% 
  group_by(price) %>% 
  summarise(total = n()) %>%
  drop_na()

n_a = df %>% 
  ungroup() %>%
  count(price, grade) %>% 
  filter(grade == "A")

join = left_join(total, n_a)
prop.test(join$n, join$total)

#join = join %>% 
 # mutate(prop = n/total) %>% 
 # select(price, prop)
```

Interpretation: since p-value << 0.05, at confidence level of 95%, we reject the null hypothesis and conclude that the proportion of restaurants receiving "A" is different among four datasets. Therefore, although the results from chi-square test tells us that dining at a particular price does not affect the overall health hazard that one faces, nonetheless the price range does seem to have an effect on whether the restaurant one dines at received an "A" as inspection grade.


### Proportion of "A"s received for Each Cuisine

Since we have received a statistically significant result for price range against receiving an "A" grade, what about the proportion of restaurants receiving "A" for each cuisine type? To answer this question, we group data by cuisine type and perform proportion test on the percentage of A received. In particular, to reduce data variability, swe concern only the cuisine types in which there are at least 10 restaurants in our dataset.

```{r}
# create dataset grouped by cuisine_type
cuisine_df <- df %>% 
  ungroup() %>%
  group_by(cuisine_description) %>%
  summarise(total_evals = n(),
            a_evals = sum(grade == "A", na.rm = TRUE)) %>%
  filter(total_evals >= 10)

test_full = cuisine_df %>%
  mutate(prop_test = map2(a_evals, total_evals, ~prop.test(.x, .y) %>%
           broom::tidy())) %>%
  unnest() %>% # pull estimated proportion and confidence intervals
  select(cuisine_description, estimate, lower_CI = conf.low, upper_CI = conf.high)

# Display result
test_full %>%
  mutate(cuisine_type = fct_reorder(cuisine_description, estimate)) %>%
  ggplot(aes(x = cuisine_type, y = estimate, ymin = lower_CI, ymax = upper_CI, col = estimate)) + 
  geom_point() + 
  geom_errorbar() +
  coord_flip() +
  theme(legend.position = "none")
```

As we can observe, soups/salads/sandwich restauants are most likely to receive As, followed by donuts, salads, and vegan.
